---
title: "Advanced Satts"
author: "Vivek Bhat Hosmat/Deepika Jantz"
date: "November 24, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#loading all the required libraries
library(data.table)
library(dplyr)
library(tidyr)
library(lubridate)
library(ggplot2)
library(reshape2)
library(ggmap)
library(knitr)
library(stringr)
library(zoo)

```




```{r}
#Gun violenece data

gun_rawdata<-fread("gun-violence-data_01-2013_03-2018.csv")

head(gun_rawdata)

gun_rawdata %>%
  distinct(state)
  

```


```{r}

city_count <- gun_rawdata %>%
  group_by(state,city_or_county) %>%
  count(city_or_county) 



```


```{r}
#Population data

population_data <- fread("popEstimate.csv")

head(population_data)

#Selecting the growth in population(estimates) between 2014 and 2017

population_growth <- population_data %>%
  select(STATE,NAME,NPOPCHG_2014,NPOPCHG_2015,NPOPCHG_2016,NPOPCHG_2017,PPOPCHG_2014,PPOPCHG_2015,PPOPCHG_2016,PPOPCHG_2017 )

population_growth_rawdata <- population_growth %>%
  select(NAME,"2014" = NPOPCHG_2014,"2015"=NPOPCHG_2015,"2016"=NPOPCHG_2016,"2017"=NPOPCHG_2017)

population_growth_percentile <- population_growth %>%
  select(NAME,"2014"=PPOPCHG_2014,"2015"=PPOPCHG_2015,"2016"=PPOPCHG_2016,"2017"=PPOPCHG_2017)

#population
population1_rawdata <- population_data %>%
  select(NAME,"2014" = POPESTIMATE2014,"2015"=POPESTIMATE2015,"2016"=POPESTIMATE2016,"2017"=POPESTIMATE2017)


```


```{r}

melted_pop_growth<-melt(population_growth_rawdata, id.vars=c("NAME"))

ggplot(melted_pop_growth, aes(variable,value,group=NAME, color=NAME)) + geom_line()


```



```{r}

melted_pop_percentile_growth<-melt(population_growth_percentile, id.vars=c("NAME"))

ggplot(melted_pop_percentile_growth, aes(variable,value,group=NAME, color=NAME)) + geom_point()


```


```{r}
#Most populated states are California and Texas

most_pop <-melted_pop_percentile_growth %>%
  filter(NAME=="California")

#Least populated states are Wyoming and Vermont

least_pop <- melted_pop_percentile_growth %>%
  filter(NAME=="Wyoming")


ggplot(most_pop, aes(variable,value,group=NAME))+ geom_line()  
  # geom_line(least_pop, mapping = variable,value)

```


```{r}
#How does the gun violations vary with the population of the state? Our assumption to start with is,Higher population will have higher gun violations.

#Linear regression

head(population_growth_rawdata)

population_growth_rawdata$`2014`<-round(rank(population_growth_rawdata$`2014`)/length(population_growth_rawdata$`2014`),2)
population_growth_rawdata$`2015`<-round(rank(population_growth_rawdata$`2015`)/length(population_growth_rawdata$`2015`),2)
population_growth_rawdata$`2016`<-round(rank(population_growth_rawdata$`2016`)/length(population_growth_rawdata$`2016`),2)
population_growth_rawdata$`2017`<-round(rank(population_growth_rawdata$`2017`)/length(population_growth_rawdata$`2017`),2)


head(population_growth_rawdata)

pop_growth_percentile <- melt(population_growth_rawdata,id.vars="NAME", variable.name = "Year", value.name ="Growth" )
  
head(pop_growth_percentile)

names(pop_growth_percentile)[1] <-"state"




head(gun_rawdata)


gun_rawdata$date <- as.Date(gun_rawdata$date)





gun_data_formatted <-gun_rawdata %>%
  group_by("Year"=year(date),state)%>%
  select(Year,state,n_killed,n_injured) %>%
  filter(Year>2013 & Year<2018) %>%
  summarize("Total_killed"=sum(n_killed),"Total_injured"=sum(n_injured))

  
head(gun_data_formatted)

gun_data_formatted$Year <- as.factor(gun_data_formatted$Year)


combined_data <- merge(pop_growth_percentile,gun_data_formatted, by=c("Year","state"))

head(combined_data)



#linear regression of growth vs number of injured
lm_population <- lm(Total_injured~Growth,combined_data)

plot(lm_population)


summary(lm_population)


#logistic regression of growth vs number of injured

log_population <- lm(Total_injured~log(Growth),combined_data)

plot(log_population)

summary(log_population)

#log of injured
log_injured <- lm(log(Total_injured)~Growth,combined_data)

plot(log_injured)

summary(log_injured)


#poly --quadratic

lm__poly_population <- lm(Total_injured~poly(Growth,2),combined_data)

plot(lm__poly_population)


summary(lm__poly_population)




```


Analysis on population data
```{r}

head(population1_rawdata)

population1_rawdata$`2014`<-round(rank(population1_rawdata$`2014`)/length(population1_rawdata$`2014`),2)
population1_rawdata$`2015`<-round(rank(population1_rawdata$`2015`)/length(population1_rawdata$`2015`),2)
population1_rawdata$`2016`<-round(rank(population1_rawdata$`2016`)/length(population1_rawdata$`2016`),2)
population1_rawdata$`2017`<-round(rank(population1_rawdata$`2017`)/length(population1_rawdata$`2017`),2)


head(population1_rawdata)

pop_percentile <- melt(population1_rawdata,id.vars="NAME", variable.name = "Year", value.name ="Population" )
  
head(pop_percentile)

names(pop_percentile)[1] <-"state"




head(gun_rawdata)


gun_rawdata$date <- as.Date(gun_rawdata$date)





gun_data_formatted <-gun_rawdata %>%
  group_by("Year"=year(date),state)%>%
  select(Year,state,n_killed,n_injured) %>%
  filter(Year>2013 & Year<2018) %>%
  summarize("Total_killed"=sum(n_killed),"Total_injured"=sum(n_injured))

  
head(gun_data_formatted)

gun_data_formatted$Year <- as.factor(gun_data_formatted$Year)


combined_data1 <- merge(pop_percentile,gun_data_formatted, by=c("Year","state"))

head(combined_data1)



#linear regression of growth vs number of injured
lm_population <- lm(Total_injured~Population,combined_data1)

plot(lm_population)


summary(lm_population)


#logistic regression of growth vs number of injured

log_population <- lm(Total_injured~log(Population),combined_data1)

plot(log_population)

summary(log_population)

#log of injured
log_injured <- lm(log(Total_injured)~Population,combined_data1)

plot(log_injured)

summary(log_injured)


#poly --quadratic

lm_poly_population <- lm(Total_injured~poly(Population,2),combined_data1)

plot(lm_poly_population)


summary(lm__poly_population)




```
























```






```{r}
pop_growth_top<-population_growth_rawdata %>%
  arrange(desc(population_growth_rawdata$`2017`))


a<- population_growth_rawdata %>%
  filter(NAME=="California")

ab<-melt(a,vars.id=a$NAME)

top.list <- dplyr::pull(ab,3)
  

pop_growth_least <- population_growth_rawdata %>%
  filter(NAME=="Wyoming")

least<- melt(pop_growth_least,vars.id=NAME)

bottom.list<- dplyr::pull(least,3)

(var_hypothesis = var.test(top.list, bottom.list, alternative="two.sided"))


```




```{r}
#Is there a seasonality influence on the number of incidents?



```


```{r}
#Is there a pattern in the places at which the incidents have taken place?



```





```{r}
#What are the commonly used guns in the incidents?



```



```{r}




```






